---
title: "cyano_barplots"
author: "Vica"
date: "9/17/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the packages

```{r echo=FALSE}
library(RColorBrewer)
library(betapart)
library(tidyverse)
library(mapproj)
library(vegan)
library(rlist)

```

# Insert csv file and filter only countries included in the Mediterranean region

```{r, echo = FALSE, include=FALSE}
med <- read.csv("../Data/Medcyanolist_source_20200130.csv", row.names = NULL)
med <- med %>% 
  filter(Med_border == "yes")
colnames(med)[1] <- "A/A"
med <- med %>% 
  mutate(continent = case_when(Country %in% c("Greece", "Italy", "Spain", "France", "Albania", "Montenegro", "Slovenia", "Portugal", "Bulgaria", "Croatia") ~ "Europe",
                               Country %in% c("Egypt", "Morocco", "Tunisia", "Libya", "Algeria", "Israel","Lebanon") ~ "Africa",Country == "Turkey" ~ "Asia"))

med <- med %>% 
mutate(lat_range = case_when(DD_Lat < 30 ~ "Low", DD_Lat > 30 & DD_Lat < 40 ~ "Medium",
                             DD_Lat > 40 ~ "High"))
```

#Number of distinct taxa in each Order, optimization in a barplot

```{r, fig.height = 4, fig.width = 6}

species_order_disti <- med %>% 
  filter(Order != "") %>%
  select(Order, Currentname) %>% 
  group_by(Order) %>% 
  summarize(Number_taxa = n_distinct(Currentname)) %>% 
  arrange(desc(Number_taxa))
species_order_disti


write.csv(species_order_disti, file = "Taxa_Order.csv") #Export the dataframe to a csv file


#brewer.pal(n = 8, name = "Accent")
#display.brewer.pal(n=8, "Accent")
#sfm <- c("#7FC97F","#BEAED4","#FDC086","#FFFF99","#386CB0","#F0027F","#BF5B17","#666666")
#sfm

species_order_disti_plot <- ggplot(species_order_disti, aes(x =reorder(Order, Number_taxa), y = Number_taxa, fill = Order )) + geom_bar(stat = "identity", colour = "black", width = 0.7) + theme_classic() + theme(legend.position = "none") + 
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent")) + coord_flip() + ylab("") +xlab("") + theme(axis.text.y = element_text(size=10, family = "serif", face = "bold", colour = "black"), axis.text.x = element_text(size = 10, family = "serif", colour = "black"))  + scale_y_continuous(breaks = seq(from = 0, to = 160, by = 20), expand = expand_scale(mult = c(0,0.1))) +
  scale_fill_manual(values = c ( 
    "Synechococcales" = "#7FC97F",
    "Nostocales" = "#BEAED4",
    "Chroococcales" = "#FDC086",
    "Oscillatoriales" = "#FFFF99",
    "Spirulinales" = "#386CB0",
    "Pleurocapsales" = "#F0027F",
    "Chroococcidiopsidales" = "#BF5B17",
    "Gloeobacterales" = "#666666"))
  
        

species_order_disti_plot
ggsave("species_order_disti_plot.png", width = 6, height = 4, dpi = 300)
```

#Number of taxa in each Family, optimization in a barplot

```{r}
species_family_all <- med %>%
  filter(Family != "") %>% 
  select(Order, Family, Currentname) %>%
  group_by(Order, Family) %>% 
  summarize(Number_of_species = n()) %>% 
  arrange(desc(Number_of_species))
species_family_all <- head(species_family_all, 20) #Select the ten richest families


species_family_all_plot <- ggplot(species_family_all, aes(x =reorder(Family, Number_of_species), y = Number_of_species, fill=Order )) + geom_bar(stat = "identity") + theme_classic() + theme(legend.position = "right") + geom_text(aes(label = Number_of_species)) 
species_family_all_plot + coord_flip() + ylab("Number of records") + xlab("") +
  scale_fill_manual(values = c (
"Synechococcales" = "#7FC97F",
"Nostocales" = "#BEAED4",
"Chroococcales" = "#FDC086",
"Oscillatoriales" = "#FFFF99",
"Spirulinales" = "#386CB0",
"Pleurocapsales" = "#F0027F",
"Chroococcidiopsidales" = "#BF5B17",
"Gloeobacterales" = "#666666"))
```

# Number of distinct taxa in each Family, optimization in a barplot

```{r fig.height = 6, fig.width= 6}
species_family_disti <- med %>%
  filter(Family != "" & Family != "Unknown") %>% 
  select(Order, Family, Currentname) %>%
  group_by(Order, Family) %>% 
  summarize(Number_of_species = n_distinct(Currentname)) %>% 
  arrange(desc(Number_of_species))
species_family_disti <- head(species_family_disti, 19)
species_family_disti

write.csv(species_family_disti, file = "Taxa_Family.csv")


species_family_disti_plot <- ggplot(species_family_disti, aes(x =reorder(Family, Number_of_species), y = Number_of_species, fill = Order )) + geom_bar(stat = "identity", colour = "black", width = 0.7) + theme_classic() + 
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent")) + coord_flip() + ylab("") + xlab("") + theme(axis.text.y = element_text(size=10, family = "serif", face = "bold", colour = "black"), axis.text.x = element_text(size = 10, family = "serif", colour = "black"))  + theme(legend.title = element_blank(), legend.position = "bottom", legend.text = element_text(color = "black", family  = "serif", size = 10))+ scale_y_continuous(breaks = seq(from = 0, to = 60, by = 10),  expand = expand_scale(mult = c(0,0.1))) +
  scale_fill_manual(values = c ( 
    "Synechococcales" = "#7FC97F",
    "Nostocales" = "#BEAED4",
    "Chroococcales" = "#FDC086",
    "Oscillatoriales" = "#FFFF99",
    "Spirulinales" = "#386CB0",
    "Pleurocapsales" = "#F0027F",
    "Chroococcidiopsidales" = "#BF5B17",
    "Gloeobacterales" = "#666666"))
species_family_disti_plot
ggsave("species_family_disti_plot.png", height = 6, width = 6, dpi = 300)
```

#Number of  taxa in each Genus, optimization in a barplot

```{r}
species_genus_all <- med %>% 
  filter(Order != "") %>% 
  select(Order, Genustotal, Currentname) %>% 
  group_by(Order, Genustotal) %>% 
  summarize(Number_of_species = n()) %>% 
  arrange(desc(Number_of_species))
species_genus_all <- head(species_genus_all, 30)

species_genus_all_plot <- ggplot(species_genus_all, aes(x =reorder(Genustotal, Number_of_species), y = Number_of_species, fill = Order )) + geom_bar(stat = "identity") + theme_classic() + theme(legend.position = "none") + geom_text(aes(label = Number_of_species)) 
species_genus_all_plot + coord_flip() + ylab("Number of records") + xlab("") +
  scale_fill_manual(values = c (
"Synechococcales" = "#7FC97F",
"Nostocales" = "#BEAED4",
"Chroococcales" = "#FDC086",
"Oscillatoriales" = "#FFFF99",
"Spirulinales" = "#386CB0",
"Pleurocapsales" = "#F0027F",
"Chroococcidiopsidales" = "#BF5B17",
"Gloeobacterales" = "#666666"))
```

#Number of distinct taxa in each Genus, optimization in a barplot

```{r fig.height=6, fig.width=6}
species_genus_disti <- med %>% 
  filter(Order != "") %>% 
  select(Order, Genustotal, Currentname) %>% 
  group_by(Order, Genustotal) %>% 
  summarize(Number_of_species = n_distinct(Currentname)) %>% 
  arrange(desc(Number_of_species))
species_genus_disti <- head(species_genus_disti, 27)
species_genus_disti

write.csv(species_genus_disti, file = "Taxa_genus.csv")


species_genus_disti_plot <- ggplot(species_genus_disti, aes(x =reorder(Genustotal, Number_of_species), y = Number_of_species, fill = Order )) + geom_bar(stat = "identity") + theme_classic() + theme(legend.position = "right") + geom_text(aes(label = Number_of_species)) 
species_genus_disti_plot + coord_flip() + ylab("Number of records") + xlab("") +
scale_fill_manual(values = c (
"Synechococcales" = "#7FC97F",
"Nostocales" = "#BEAED4",
"Chroococcales" = "#FDC086",
"Oscillatoriales" = "#FFFF99",
"Spirulinales" = "#386CB0",
"Pleurocapsales" = "#F0027F",
"Chroococcidiopsidales" = "#BF5B17",
"Gloeobacterales" = "#666666"))

species_genus_disti_plot <- ggplot(species_genus_disti, aes(x =reorder(Genustotal, Number_of_species), y = Number_of_species, fill = Order )) + geom_bar(stat = "identity", colour = "black", width = 0.7) + theme_classic() + 
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent")) + coord_flip() + ylab("") + xlab("") + theme(axis.text.y = element_text(size=10, family = "serif", face = "bold.italic",colour = "black", ), axis.text.x = element_text(size = 10, family = "serif", colour = "black"))  + theme(legend.title = element_blank(), legend.position = "bottom", legend.text = element_text(color = "black", family  = "serif", size = 10))+ 
  guides(fill=guide_legend(nrow=2,byrow=TRUE)) + 
  scale_y_continuous(breaks = seq(from = 0, to = 35, by = 5), expand = expand_scale(mult = c(0,0.1))) +
    scale_fill_manual(values = c ( 
    "Synechococcales" = "#7FC97F",
    "Nostocales" = "#BEAED4",
    "Chroococcales" = "#FDC086",
    "Oscillatoriales" = "#FFFF99",
    "Spirulinales" = "#386CB0",
    "Pleurocapsales" = "#F0027F",
    "Chroococcidiopsidales" = "#BF5B17",
    "Gloeobacterales" = "#666666"))

species_genus_disti_plot
ggsave("species_genus_disti_plot.png", height = 6, width = 6, dpi = 300)
```

#Taxa found in habitat Reservoir

```{r}
reserv.hab.taxa.disti <- med %>%  select(Currentname, Isolationsource, Order, Country) %>%
  filter(Order != "") %>% 
  filter((str_detect(Isolationsource, pattern = "Reservoir"))) %>% 
  group_by(Country, Isolationsource) %>% 
  summarise(Taxa_records = n_distinct(Currentname)) %>% 
  arrange(desc(Taxa_records))
reserv.hab.taxa.disti

```

# Number of taxa in different Habitats

```{r}
isol_per_country <- med %>% 
  filter(Order != "") %>%
  select(Country, Isolationsource, Habitat_source) %>% 
  group_by(Country, Isolationsource) %>% 
  summarize(N = n_distinct("Isolationsource"))
isol_per_country
isol_per_country %>% group_split(Country)

habitat.taxa.plot <- ggplot(habitat.taxa, aes(x=reorder(Country,N), y = N, fill = Habitat_source)) + geom_bar(stat = "identity")+ theme_classic() + theme(legend.position = "none") + geom_text(aes(label = N)) 
habitat.taxa.plot + coord_flip() + ylab("Number of taxa") + xlab("") +
  scale_fill_brewer(palette = "Paired")
```

#Number of taxa in different Habitat types

```{r}
habitat.type.taxa <- med %>% 
  select(Habitat_type, Currentname, Order) %>% 
  filter(Order != "") %>% 
  group_by(Habitat_type) %>% 
  summarize(N=n_distinct(Currentname)) %>% 
  arrange(desc(N))
habitat.type.taxa <- habitat.type.taxa[-3,]

habitat.type.taxa.plot <- ggplot(habitat.type.taxa, aes(x=reorder(Habitat_type,N), y = N, fill = Habitat_type)) + geom_bar(stat = "identity")+ theme_classic() + theme(legend.position = "none") + geom_text(aes(label = N)) 
habitat.type.taxa.plot + coord_flip() + ylab("Number of taxa") + xlab("")
```




```{r}

taxa.habitat.type <- split(med, med$Habitat_type)
taxa.habitat.type

## Find unique taxa per Habitat type

unique.taxa.habitat.type <- sapply(taxa.habitat.type, function(x){
  unique(x["Currentname"])
})
unique.taxa.habitat.type

list.save(unique.taxa.habitat.type, "hab_type.rds")


## Find unique taxa per Habitat

taxa.habitat <- split(med, med$Habitat)
taxa.habitat

unique.taxa.habitat <- sapply(taxa.habitat, function(x){
  unique(x["Currentname"])
})
unique.taxa.habitat.type
```

# Different taxa in each Country, optimization in barplot 

```{r}
vec <- as.vector(med$Currentname)
unique_species <- names(which(table(vec) == 1))
unique_species_df <- med[med$Currentname %in% unique_species,]

unique_species_count <- unique_species_df %>% 
  group_by(Country, continent) %>% 
  summarize(Number_unique_species = length(Currentname))
unique_species_count

write.csv(unique_species_count, file = "Unique_species_per_Country.csv")

#colourCount = length(unique(unique_species_count$Country))


unique_species_count_plot <- ggplot(unique_species_count, aes(x=reorder(Country, Number_unique_species), y=Number_unique_species, fill = continent)) + geom_bar(stat="identity") + theme_classic() + theme(legend.position = "right") + geom_text(aes(label=Number_unique_species)) + coord_flip() + ylab("Number of taxa") + xlab("")
unique_species_count_plot
#ggsave("unique_species_count_plot.png")
```

# Number of taxa in each Country, optimization in barplot

```{r}
taxa.disti <- med %>%
  filter(Species != "")%>%
  select(Title, Currentname, continent)%>%
  group_by(Title, continent)%>%
  summarize(Taxa_number = n_distinct(Title))%>%
arrange(desc(Title))
taxa.disti

##write.csv(taxa.disti, file = "Taxa_Country.csv")

taxa.disti.plot <- ggplot(taxa.disti, aes(x=reorder(continent, Taxa_number), y=Taxa_number, fill = continent)) + geom_bar(stat="identity") +
        theme_classic() + theme(legend.position = "right") + coord_flip() + ylab("") + xlab("")
taxa.disti.plot
#ggsave("taxa.disti.plot.png")
```





